Diffusive Homeostasis in a Self-Organizing Recurrent Neural Network \\ \begin{large} Spatially dependent Interaction as a Determinant of Neural Activity and Plasticity \end{large}

### 2017-01-24 draft.tex
.

\section{Introduction}
Among other functional properties of the mammalian central nervous system, it is involved in maintaining crucial properties of the body in a functional regime. Some examples regarding the organism as a whole are e.g. body temperature and blood pressure, both being controlled by hypothalamus \cite[p. 484]{Bear_Exploring_the_Brain}. The term \textit{homeostasis} entailing these regulatory mechanisms was first introduced by Walter C. Cannon in 1932 \cite{Wisdom_of_the_Body_Cannon, Homeostasis_Queenan_2012}. In more recent years, experimental studies have found evidence for processes within the brain that do not function as a regulatory control for certain aspects of the body, but as a homeostatic control of neural activity itself \cite{Turrigiano_1994,Burrone_2003}. Many processes underlying this self-regulatory behavior have been found up till now, and many different biophysical interdependencies have shown to play a role. This diversity makes comparability of these mechanisms a rather difficult task. On a relatively general level though, experiments have revealed two main types of neuronal regulation: synaptic homeostatic plasticity and intrinsic homeostasis \cite{Desai_2003}.

@@

This thesis is based on earlier versions of a self-organizing recurrent neural network (SORN), in particular an implementation using leaky integrate-and-fire neurons (LIF-SORN) \cite{SORN_Paper}. It was developed to model features of activity and topology of cortical networks, in particular L5 of the neocortex. In these previous implementations, a slow homeostatic mechanism was implemented among excitatory neurons by an intrinsic adaptation of excitability. This led to the aforementioned problems of homeostatic control: Neural activity was tuned to a globally set target, allowing for no variability. In contrast, cortical networks are known to exhibit a broad distribution of firing rates, which is well described by a right-skewed, log-normal like shape \cite{Buzsaki_Fir_Rates_2014,Wohrer_Fir_Rates_2012}. If the LIF-SORN was meant to provide a good model for cortical structure and activity, it would require a different type of homeostatic model.

In this thesis, we examined \textit{diffusive homeostasis} as a possible candidate to replace the previously used model of intrinsic plasticity. The idea and modeling of diffusive homeostasis was adopted from a paper by Y. Sweeney et al. \cite{Sweeney_Paper}, which models neural tissue as a two-dimensional surface and a set of points representing the neurons' positions. The group of excitatory neurons acted as a point-source of nitric oxide (NO), as well as as a sensor for the NO-concentration at each individual position. The production and sensing of NO forms the basis of a feedback loop: The individual NO-readout is fed into a comparator which causes an appropriate change within the internal firing threshold of the neuron, in turn altering the neuron's firing rate. The control system is then closed by linking the rate of NO-production to the neuron's firing rate. The results by Sweeney et al. suggest that diffusive signaling could resolve the dichotomy between overall stability of firing activity and the need to allow flexibility among individual neurons' activities. Intuitively speaking, through the diffusive signal, each neurons receives a "mixture" of its own activity and its neighboring neurons. Thereby individual tuning of firing rates is suppressed while the overall population activity is kept at a constant level.

Key aspects of this thesis include an analysis of the stability of the homeostatic control, followed by a comparison of features of the original LIF-SORN and the diffusive variant. We expect to observe a preservation of non-random features that have been found in the original LIF-SORN while incorporating a stronger variance within neural activity which has previously been suppressed by the rigid single-cell homeostasis. In the face of possible \emph{new} features within the network's structure, we will clarify the causal relation between diffusive spatial interaction and synaptic topology.

@@

Our original model of homeostatic control was described as an operation over discrete time steps $\Delta t = 0.1ms$, carried out for each excitatory neuron:

\begin{align}
V_t &\rightarrow V_t + \eta_{IP} (N_{spikes} - h_{IP}) \label{can_hom_1}\\
N_{spikes} &\rightarrow 0 \label{can_hom_2}
\end{align}
where $V_t$ is the firing threshold, $\eta_{IP}$ an adaption rate and $h_{IP}$ the desired number of spikes per time step. $N_{spikes}$ is a variable, counting the number of spikes of the neuron within each interval. In a continuous, rate-based form, this update rule can as well be written as:

\begin{equation}
\dot{V}_t = \eta_{IP}(r-r_{IP}) \label{can_hom_rate}
\end{equation}
with $r$ as the neuron's firing rate and $r_{IP}=h_{IP}/\Delta t$ the target firing rate. This feedback control indirectly drives the firing rate of each neuron towards $r_{IP}$: If $r>(<)r_{IP}$, $V_t$ increases (decreases), reducing (increasing) the probability of a spike to occur.

The model presented in \cite{Sweeney_Paper} includes spatial interaction across excitatory neurons through the diffusion of nitric oxide (NO). Enzymes that are responsible for NO synthesis (NOS) are present in different areas of the body, namely being involved in the dilation of blood vessels (endothelial NOS, eNOS) , the immune system (inducible NOS, iNOS) and the CNS (neuronal NOS, nNOS) \cite{NOS_Mammals}. These Enzymes differ in their functionality and dependence on the presence of other molecules. In particular, nNOS is sensitive to the concentration of $Ca^{2+}$ ionic Calcium \cite{Knowles_Ca_nNOS}.

On the other hand, experimental studies have suggested that NO can act as a diffusive signaling pathway, decreasing intrinsic excitability \cite{Steinert_NO}, or generally suggest a role of NO in maintaining a functional state of activity \cite{Pape_NO}.

Sweeney et al. combined these empirical findings into a model of \emph{diffusive homeostasis}, which is governed by the following equations:

\begin{align}
\dot{{Ca^{2+}}^i}(t) &= -\frac{{Ca^{2+}}^i}{\tau_{Ca^{2+}}} + Ca^{2+}_{spike} \sum_{j} \delta(t-t^i_{spike,j}) \label{Ca_dyn}\\
\dot{nNOS^i}(t) &= \frac{1}{\tau_{nNOS}} \left(  \frac{{{Ca^{2+}}^i}^3}{{{Ca^{2+}}^i}^3+1} - nNOS^i \right) \label{nNOS_dyn}\\
\dot{NO}(\mathbf{r},t)&=-\lambda NO + D \nabla^2 NO + \sum_{i} \delta^2(\mathbf{r}-\mathbf{r}_{neur}^i)\cdot nNOS^i \label{NO_dyn}\\
\dot{V_t^i}(t) &= \frac{NO(\mathbf{r}_{neur}^i,t)-NO_0}{NO_0\cdot\tau_{V_t}} \label{Theta_dyn}
\end{align}

A depolarization within a nerve cell upon a spike-event $t_{spike}$ causes a fixed inflow of ionic current $Ca^{2+}_{spike}$, which is modeled as an instantaneous increase of the $Ca^{2+}$ concentration. Over time, the concentration decays exponentially by a time constant $\tau_{Ca^{2+}}$, see \eqref{Ca_dyn}. Though $Ca^{2+}$ currents can be described in a much more detailed fashion, it can be considered as a reasonable approximation \cite[p.~198-203]{Theor_Neur_Dayan}. The aforementioned influence of $Ca^{2+}$ onto nNOS was modeled by Sweeney et al. through \eqref{nNOS_dyn}, using the Hill equation \cite{Hill_Equ} to model a cooperative binding mechanism. The $nNOS$ production is then fed into the "pool" of nitric oxide via point sources located at the neurons' positions. Apart from the inflow and the diffusive term $D \nabla^2 NO$, an additional decay term was added to provide a stable finite $NO$ concentration under constant neuronal activity.

@@

\section{Further Analysis of the Firing Rate Distribution}
On a preliminary level, the results of section \ref{Fir_Dist_Section} were positive with respect to our goal of achieving a broad and skewed firing rate distribution. Following this, we tried to further gain insight into the actual shape of the resulting distribution. 
\subsection{Review of a Dynamic Mean-Field Model}
Sweeney et al. attempted to predict the shape of the firing rate distribution by means of a simplified mean-field approximation of diffusion \cite{Sweeney_Paper}. The main simplification within homeostatic feedback was the replacement of Equation \ref{Theta_dyn} by
\begin{equation}
\dot{V_t^i}(t) = \frac{1}{\tau_{HIP}} \left( (1-\alpha)\frac{\phi_i-\phi_0}{\phi_i} +\alpha \frac{\langle \phi \rangle-\phi_0}{\langle \phi \rangle} \right)\;. 
\label{diff_hom_simpl_sweeney}
\end{equation}
$\mathrm{\alpha \: \epsilon \: [0,1]}$ thereby acted as a parameter that determines the "mixture" between single-neuron-homeostasis ($\mathrm{\alpha=0}$) and the limit of quasi instantaneous spreading of the diffusive signal across the population ($\mathrm{\alpha=1}$). $\mathrm{\phi_i}$ and $\mathrm{\mathrm{\langle \phi \rangle}}$ represent individual firing rates and their overall mean.

@@

@@

\subsubsection{Possible Modifications}\label{Possible_Modifications_Section}
Since these considerations evoked a rather skeptical view onto the biological plausibility of these results, we considered possible modifications of the homeostatic model. One should note that these are to be evaluated in the context of providing a model of an actual physiological process. If the main goal was to simply provide a homeostatic method that retains an overall level of activity without tightly fixing individual rates, we would suggest to use the idealized model of instant diffusion, especially because of its computational simplicity. However, as already mentioned, the given experimentally based parameters do not allow to accept this simplification from the standpoint of biological and physical accuracy. Instead, two elements of the model could be put into question: First, is it appropriate to assume an approximately point-like source of nitric oxide? And second, are the relatively simple dynamics of threshold adaption (Equation \eqref{Theta_dyn}) an acceptable description of the actual physiological processes involved?

While Sweeney et al. argued that the size of the studied tissue is large compared to the size of individual somata (which, apparently, they claim to be the main source of nitric oxide), Philippides et al. reported that NO is often synthesized in a more delocalized way by means of fine fibers exceeding the soma \cite{Philippides_2005}. The diameter of these "production areas" is on the order of $\mathrm{10-100 \mu m}$. To incorporate this spread of synthesis in our model of diffusion, one would - for each neuron - need to convolve a suitably shaped and shifted "production"-kernel (scaled according to the momentary rate of NO-synthesis) with the field of NO-concentration. Intuitively, this would locally smear the distinction between neighboring neurons. From a more abstract point of view, it would provide a way to drive the matrix defined in Equation \eqref{NO_equil_cond_interact_matrix_operator} towards singularity, allowing for more states of activity to approximately fulfill the homeostatic target.

Regarding the second remark, judging the plausibility of the used model of threshold adaptation is a rather problematic task. Experimental studies have investigated the effect of NO onto the conductance of ionic channels \cite{Steinert_NO,Steinert_NO_2011}. Mapping these features onto a simple LIF-model is obviously not easy, and many consequences of NO-modulation onto neural dynamics can certainly not be represented in such a simple model. Our criticism is thus rather based on more general arguments.

@@

In either case, such a modification would break the ability of the network to tune thresholds in such a way that necessarily \emph{all} neurons eventually reach target concentration. As a consequence, the fixed point of firing rates under diffusive homeostasis would not be strictly determined by spatial structure, but by a combination of network structure and the neurons' positions. As an example, consider a modification of Equation \eqref{Theta_dyn} of the form
\begin{equation}
\dot{V_t^i}(t) = \frac{1}{\tau_{V_t}}\left(\frac{NO(\mathbf{r}_{neur}^i,t)-NO_0}{NO_0} + \frac{\alpha}{V_t^i - V_{t,min}} + \frac{\alpha}{V_t^i - V_{t,max}}\right) \; .
\label{Theta_dyn_with_Bounds}
\end{equation}
In this form, $\mathrm{NO(\mathbf{r}_{neur}^i,t) = NO_0}$ does not follow any more as the fixed point solution. Rather, one needs to consider the fact that, given the recurrent network topology, the steady state of $\mathrm{NO(\mathbf{r}_{neur}^i)^* = NO(\mathbf{r}_{neur}^i,V_t^1,...,V_t^n)^*}$ in turn depends on the set of thresholds. Unfortunately, this increase in complexity would presumably make it impossible to find a general analytic description of the fixed point similar to the one introduced and discussed in Section \ref{Matrix_Diff_Model_Section}. Moreover, while this modification would provide a way to incorporate network topology into the theoretical steady state of the system for finite diffusion constants, it is not clear whether it would allow perturbations (e.g. due to external input) to be actually sustained longer.

@@

\section{Conclusion}\label{Conclusion_Section}
Drawing a conclusion from our results presents itself as a somewhat ambiguous task. On a phenomenological level, our main goal of allowing the LIF-SORN to exhibit a broad, skewed distribution was achieved. However, whether the implications of the diffusive model that revealed itself upon deeper investigations are acceptable in terms of biological plausibility is still left open to debate.

@@

>>>>

\section{Introduction}
\subsection{Homeostasis in the Nervous System}
Among other functional properties of the mammalian central nervous system, it is involved in maintaining crucial properties of the body in a functional regime. Some examples regarding the organism as a whole are e.g. body temperature and blood pressure, both being controlled by hypothalamus \cite[p. 484]{Bear_Exploring_the_Brain}. The term \textit{homeostasis} entailing these regulatory mechanisms was first introduced by Walter C. Cannon in 1932 \cite{Wisdom_of_the_Body_Cannon, Homeostasis_Queenan_2012}. In more recent years, experimental studies have found evidence for processes within the brain that do not function as a regulatory control for certain aspects of the body, but as a homeostatic control of neural activity itself \cite{Turrigiano_1994,Burrone_2003}. Many processes underlying this self-regulatory behavior have been found up till now, and many different biophysical interdependencies have shown to play a role. This diversity makes comparability of these mechanisms a rather difficult task. On a relatively general level though, experiments have revealed two main types of neuronal regulation: synaptic homeostatic plasticity and intrinsic homeostasis \cite{Desai_2003}.

@@

\subsection{Diffusive Homeostasis}
In this thesis, we examined \textit{diffusive homeostasis} as a possible candidate to replace the previously used model of intrinsic plasticity. The idea and modeling of diffusive homeostasis was adopted from a paper by Y. Sweeney et al. \cite{Sweeney_Paper}, which models neural tissue as a two-dimensional surface and a set of points representing the neurons' positions. The group of excitatory neurons acted as a point-source of nitric oxide (NO), as well as as a sensor for the NO-concentration at each individual position. The production and sensing of NO forms the basis of a feedback loop: The individual NO-readout is fed into a comparator which causes an appropriate change within the internal firing threshold of the neuron, in turn altering the neuron's firing rate. The control system is then closed by linking the rate of NO-production to the neuron's firing rate. The results by Sweeney et al. suggest that diffusive signaling could resolve the dichotomy between overall stability of firing activity and the need to allow flexibility among individual neurons' activities. Intuitively speaking, through the diffusive signal, each neurons receives a "mixture" of its own activity and its neighboring neurons. Thereby individual tuning of firing rates is suppressed while the overall population activity is kept at a constant level.

\subsubsection{Nitric Oxide in the Nervous System}\label{NO_Experiments_Section}
The model presented in \cite{Sweeney_Paper} includes spatial interaction across excitatory neurons through the diffusion of nitric oxide (NO). Enzymes that are responsible for NO synthesis (NOS) are present in different areas of the body, namely being involved in the dilation of blood vessels (endothelial NOS, eNOS) , the immune system (inducible NOS, iNOS) and the CNS (neuronal NOS, nNOS) \cite{NOS_Mammals}. These Enzymes differ in their functionality and dependence on the presence of other molecules. In particular, nNOS is sensitive to the concentration of $Ca^{2+}$ ionic Calcium \cite{Knowles_Ca_nNOS,Steinert_NO}. Because of its small molecular structure and electrical neutrality, nitric oxide that was synthesized in the nervous system can diffuse freely, which makes it a good candidate to be involved into form of diffusive intracellular communication \cite{Lancaster_1994}.

Experimental studies have also suggested that NO can act as an inhibitory diffusive signaling pathway, decreasing intrinsic excitability \cite{Steinert_NO_2011}, or generally suggest a role of NO in maintaining a functional state of activity \cite{Pape_NO}. In experiments upon neurons in the medial nucleus of the trapezoid body, Steinert et al. found that cells that were subject to NO-signaling by an external source exhibited an increment in outward $\mathrm{K^+}$ current, which, due to its hyperpolarizing effect, can be regarded as a form of intrinsic inhibition.

\subsection{Current State of the Network Model}
This thesis is based on earlier versions of a self-organizing recurrent neural network (SORN), in particular an implementation using leaky integrate-and-fire neurons (LIF-SORN) \cite{SORN_Paper}. It was developed to model features of activity and topology of cortical networks, in particular L5 of the neocortex. In these previous implementations, a slow homeostatic mechanism was implemented among excitatory neurons by an intrinsic adaptation of excitability. This led to the aforementioned problems of homeostatic control: Neural activity was tuned to a globally set target, allowing for no variability. In contrast, cortical networks are known to exhibit a broad distribution of firing rates, which is well described by a right-skewed, log-normal like shape \cite{Buzsaki_Fir_Rates_2014,Wohrer_Fir_Rates_2012}. Thus, if the LIF-SORN was meant to provide a good model for cortical structure \emph{and} activity, it would require a different type of homeostatic model.

\subsection{Thesis Outline}
Key aspects of this thesis include an analysis of the stability of the homeostatic control, followed by a comparison of features of the original LIF-SORN and the diffusive variant. We expect to observe a preservation of non-random features that have been found in the original LIF-SORN while incorporating a stronger variance within neural activity which has previously been suppressed by the rigid single-cell homeostasis. In the face of possible \emph{new} features within the network's structure, we will clarify the causal relation between diffusive spatial interaction and synaptic topology.

@@

Our original model of homeostatic control was described as an operation over discrete time steps $\Delta t = 0.1ms$, carried out for each excitatory neuron:
\begin{align}
V_t &\rightarrow V_t + \eta_{IP} (N_{spikes} - h_{IP}) \label{can_hom_1}\\
N_{spikes} &\rightarrow 0 \label{can_hom_2}
\end{align}
where $V_t$ is the firing threshold, $\eta_{IP}$ an adaption rate and $h_{IP}$ the desired number of spikes per time step. $N_{spikes}$ is a variable, counting the number of spikes of the neuron within each interval. In a continuous, rate-based form, this update rule can as well be written as:
\begin{equation}
\dot{V}_t = \eta_{IP}(r-r_{IP}) \label{can_hom_rate}
\end{equation}
with $r$ as the neuron's firing rate and $r_{IP}=h_{IP}/\Delta t$ the target firing rate. This feedback control indirectly drives the firing rate of each neuron towards $r_{IP}$: If $r>(<)r_{IP}$, $V_t$ increases (decreases), reducing (increasing) the probability of a spike to occur.

The new diffusive homeostatic model by Sweeney et al. combined the empirical findings discussed in Section \ref{NO_Experiments_Section} into a set of differential equations:   
\begin{align}
\dot{{Ca^{2+}}^i}(t) &= -\frac{{Ca^{2+}}^i}{\tau_{Ca^{2+}}} + Ca^{2+}_{spike} \sum_{j} \delta(t-t^i_{spike,j}) \label{Ca_dyn}\\
\dot{nNOS^i}(t) &= \frac{1}{\tau_{nNOS}} \left(  \frac{{{Ca^{2+}}^i}^3}{{{Ca^{2+}}^i}^3+1} - nNOS^i \right) \label{nNOS_dyn}\\
\dot{NO}(\mathbf{r},t)&=-\lambda NO + D \nabla^2 NO + \sum_{i} \delta^2(\mathbf{r}-\mathbf{r}_{neur}^i)\cdot nNOS^i \label{NO_dyn}\\
\dot{V_t^i}(t) &= \frac{NO(\mathbf{r}_{neur}^i,t)-NO_0}{NO_0\cdot\tau_{V_t}} \label{Theta_dyn}
\end{align}

A depolarization within a nerve cell upon a spike-event $t_{spike}$ causes a fixed inflow of ionic current $Ca^{2+}_{spike}$, which is modeled as an instantaneous increase of the $Ca^{2+}$ concentration. Over time, the concentration decays exponentially by a time constant $\tau_{Ca^{2+}}$, see \eqref{Ca_dyn}. Though $Ca^{2+}$ currents can be described in a much more detailed fashion, it can be considered as a reasonable approximation \cite[p.~198-203]{Theor_Neur_Dayan}. The influence of $Ca^{2+}$ onto nNOS was modeled by Sweeney et al. through \eqref{nNOS_dyn}, using the Hill equation \cite{Hill_Equ} to model a cooperative binding mechanism. The $nNOS$ production is then fed into the "pool" of nitric oxide via point sources located at the neurons' positions. Apart from the inflow and the diffusive term $D \nabla^2 NO$, an additional decay term was added to provide a stable finite $NO$ concentration under constant neuronal activity.

@@

\section{Further Analysis of the Firing Rate Distribution}
On a preliminary level, the results of section \ref{Fir_Dist_Section} were positive with respect to our goal of achieving a broad and skewed firing rate distribution. Following this, we tried to further gain insight into the actual shape of the resulting distribution. 
\subsection{Review of a Dynamic Mean-Field Model}\label{Mean_Field_Review_Section}
Sweeney et al. attempted to predict the shape of the firing rate distribution by means of a simplified mean-field approximation of diffusion \cite{Sweeney_Paper}. The main simplification within homeostatic feedback was the replacement of Equation \ref{Theta_dyn} by
\begin{equation}
\dot{V_t^i}(t) = \frac{1}{\tau_{HIP}} \left( (1-\alpha)\frac{\phi_i-\phi_0}{\phi_i} +\alpha \frac{\langle \phi \rangle-\phi_0}{\langle \phi \rangle} \right)\;. 
\label{diff_hom_simpl_sweeney}
\end{equation}
$\mathrm{\alpha \: \epsilon \: [0,1]}$ thereby acted as a parameter that determines the "mixture" between single-neuron-homeostasis ($\mathrm{\alpha=0}$) and the limit of quasi instantaneous spreading of the diffusive signal across the population ($\mathrm{\alpha=1}$). $\mathrm{\phi_i}$ and $\mathrm{\mathrm{\langle \phi \rangle}}$ represent individual firing rates and their overall mean.

@@

\subsection{Spreading of NO-sources extends Encoding of Input Heterogeneity}
Since our analytic results revealed that a strong correlation between spatial structure and result firing rates is present if diffusive interaction allows a clear distinction between individual neurons and their spatial neighbors, we considered the possibility that a wider spread of each individual NO-source could locally decrease distinguishability between neurons. Thereby, it might allow input heterogeneities to be encoded over a longer period of time without the need to tweak diffusion to unrealistically high values. The biological plausibility of this idea is discussed in Section \ref{Possible_Modifications_Section}.

To test our hypothesis, we modified our model such that each neuron released nitric oxide evenly into a circular area with a radius of $\mathrm{50 \, \mu m}$, centered around the neuron's position. The homeostatic readout was still taken solely from the centering point. Thereafter, we ran the same simulation as for Figure \ref{Ext_Input_Switch}. Comparing both plots, we saw a much greater persistence of distinction between excitatory subgroups. The time course of activity suggests that this distinction might even be permanent and steady instead of a very slow but transient state.
\begin{figure}
%% spread_source_illustration.py
\includegraphics[width=\textwidth]{../../plots/spread_source_illustration.png}
%% plot_external_input_switch_activity_spread_source.py
\includegraphics[width=\textwidth]{../../plots/external_input_switch_spread_source.png}
\caption{A/B: Illustration of circular kernels used for NO-insertion, overlay of all 400 neurons (left) and a single example (right). C: Resulting activity for excitatory subgroups. All settings were kept the same as for Figure \ref{Ext_Input_Switch}.}
\label{External_Input_Switch_Spread_Source}
\end{figure}

@@

\subsubsection{Modifications of the Homeostatic Model}\label{Possible_Modifications_Section}
Since these considerations evoked a rather skeptical view onto the biological plausibility of these results, we considered possible modifications of the homeostatic model. One should note that these are to be evaluated in the context of providing a model of an actual physiological process. If the main goal was to simply provide a homeostatic method that retains an overall level of activity without tightly fixing individual rates, we would suggest to use the idealized model of instant diffusion, especially because of its computational simplicity. However, as already mentioned, the given experimentally based parameters do not allow to accept this simplification from the standpoint of biological and physical accuracy.

Instead, we hypothesized that by spatially increasing the spread of NO-insertion, we might be able to uncouple activity from spatial structure without tweaking the diffusion process. While Sweeney et al. argued that the size of the studied tissue is large compared to the size of individual somata (which, apparently, they claim to be the main source of nitric oxide), Philippides et al. reported that NO is often synthesized in a more delocalized way by means of fine fibers exceeding the soma \cite{Philippides_2005}. The diameter of these "production areas" is on the order of $\mathrm{10-100 \mu m}$. To incorporate this spread of synthesis in our model of diffusion, we convolved a circular shaped and shifted "production"-kernel (scaled according to the momentary rate of NO-synthesis) with the field of NO-concentration. Intuitively, this locally smeared the distinction between neighboring neurons. From a more abstract point of view, it provided a way to drive the matrix defined in Equation \eqref{NO_equil_cond_interact_matrix_operator} towards singularity, allowing for more states of activity to approximately fulfill the homeostatic target. As Figure \ref{External_Input_Switch_Spread_Source} suggests, this indeed improved flexibility of individual activity while retaining an overall constant firing rate. By adjusting the area of NO insertion we thereby gained further control over the strictness of the homeostatic mechanism.

Another point of criticism that might justify future changes of the homeostatic model is the plausibility of the used model of threshold adaptation. Experimental studies have investigated the effect of NO onto the conductance of ionic channels \cite{Steinert_NO,Steinert_NO_2011}. Mapping these features onto a simple LIF-model is obviously not easy, and many consequences of NO-modulation onto neural dynamics can certainly not be represented in such a simple model. Our criticism is thus rather based on more general arguments.

@@

In either case, such a modification could break the ability of the network to tune thresholds in such a way that necessarily \emph{all} neurons eventually reach target concentration. As a consequence, the fixed point of firing rates under diffusive homeostasis would not be strictly determined by spatial structure, but by a combination of network structure and the neurons' positions. As an example, consider a modification of Equation \eqref{Theta_dyn} of the form
\begin{equation}
\dot{V_t^i}(t) = \frac{1}{\tau_{V_t}}\left(\frac{NO(\mathbf{r}_{neur}^i,t)-NO_0}{NO_0} + \frac{\alpha}{V_t^i - V_{t,min}} + \frac{\alpha}{V_t^i - V_{t,max}}\right) \; .
\label{Theta_dyn_with_Bounds}
\end{equation}
In this form, $\mathrm{NO(\mathbf{r}_{neur}^i,t) = NO_0}$ does not follow any more as the fixed point solution. Rather, one needs to consider the fact that, given the recurrent network topology, the steady state of $\mathrm{NO(\mathbf{r}_{neur}^i)^* = NO(\mathbf{r}_{neur}^i,V_t^1,...,V_t^n)^*}$ in turn depends on the set of thresholds. Unfortunately, this increase in complexity would presumably make it impossible to find a general analytic description of the fixed point similar to the one introduced and discussed in Section \ref{Matrix_Diff_Model_Section}. Moreover, while this modification would provide a way to incorporate network topology into the theoretical steady state of the system for finite diffusion constants, it is not clear whether it would allow perturbations (e.g. due to external input) to be actually sustained longer.

@@

\section{Conclusion and Outlook}\label{Conclusion_Section}
Drawing a conclusion from our results presents itself as a somewhat ambiguous task. On a phenomenological level, our main goal of allowing the LIF-SORN to exhibit a broad, skewed distribution was achieved. However, whether the implications of the diffusive model that revealed itself upon deeper investigations are acceptable in terms of biological plausibility is still left open to debate.

@@

With regard to the future use and implementation of diffusive homeostasis, it could be desirable to cover the essential effects in a more simple and abstract formalism. As shown in Section \ref{Mean_Field_Review_Section}, a simple mean field model as presented by Sweeney et al. still carries problems similar to those of single-neuron homeostasis. However, modifications of this model that incorporates our remarks in Section \ref{Possible_Modifications_Section} might be able to recover properties of the full diffusive homeostatic feedback while being more computationally efficient.
